---
title: "AMPBA SA3 Batch 15 - Individual Assignment"
author: "Deep Kamal Singh - 12020053"
date: "11/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##  Homework PART-A

The Current Population Survey (CPS), undertaken by the U.S. Census Bureau, periodically conducts a variety of surveys on a variety of topics. In this example we look at a cross- section of 1,289 persons interviewed in March 1995 to study the factors that determine hourly wage (in dollars) in this sample.16 Keep in mind that these 1,289 observations are a sample from a much bigger population. We will consider these 1289 observations as a population for our purpose.


```{r library}
library(openxlsx)
setwd("/Users/dks/AMPBA/Term3/SA3/HW1/")
```

### The variables used in the analysis are defined as follows:

**Wage**: Hourly wage in dollars, which is the dependent variable.

*The explanatory variables, or regressors, are as follows:*

**Female**: Gender, coded F for female, M for male

**Nonwhite**: Race, coded N for non-white workers, W for white workers

**Union**: Union status, coded M if in a union member, N otherwise

**Education**: Education (in years)

**Exper**: Potential work experience (in years), defined as age minus years of schooling minus 6. (It is assumed that schooling starts at age 6.)

**Age**: Age of the worker

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
